---
title: "Testing PH assumptions"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: references.bib
csl: biomedicine.csl
vignette: >
  %\VignetteIndexEntry{Kaplan-Meier Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.dim = c(8, 6),
  warning = FALSE,
  message = FALSE
)
```

# Loading R packages

```{r}
# install.packages("maicplus")
library(maicplus)
```

# Introduction

Here we demonstrate briefly how we can test proportional hazards assumptions including weights for time-to-event endpoints.

```{r}
data(weighted_sat)
data(adtte_sat)
data(pseudo_ipd_sat)
```

## Unanchored case

```{r}
result <- maic_unanchored(
  weights_object = weighted_sat,
  ipd = adtte_sat,
  pseudo_ipd = pseudo_ipd_sat,
  trt_ipd = "A",
  trt_agd = "B",
  normalize_weight = FALSE,
  endpoint_name = "Overall Survival",
  endpoint_type = "tte",
  eff_measure = "HR",
  time_scale = "months",
  km_conf_type = "log-log"
)

fit_surv <- result$inferential$fit$km_after
fit_cox <- result$inferential$fit$model_after
```

To assess if the hazards are proportional to each other over time, we can check if the log cumulative hazard lines are parallel.

```{r}
# Log-log plot
ph_diagplot_lch(fit_surv,
  time_scale = "months", log_time = TRUE,
  endpoint_name = "OS", subtitle = "(After Matching)"
)
```

Another way is using the Schoenfeld residual.

At the subject’s failure time, the residual measures how the value of $x$ for the subject who fails differs from a weighted average of $x$ values for those still at risk. Weights depend on estimated HR for each subject at risk. [@brown]

If the lines in Schoenfeld plots are flat, covariate effect are constant over time and proportional hazards assumptions are met. Alternatively, a non-significant p-value in Schoenfeld residual test suggests that there is no strong evidence for non-proportionality.

```{r}
# Schoenfeld plot
ph_diagplot_schoenfeld(fit_cox,
  time_scale = "months", log_time = FALSE,
  endpoint_name = "OS", subtitle = "(After Matching)"
)

# Or set of diagnostic plots
ph_diagplot(
  weights_object = weighted_sat,
  tte_ipd = adtte_sat,
  tte_pseudo_ipd = pseudo_ipd_sat,
  trt_var_ipd = "ARM",
  trt_var_agd = "ARM",
  trt_ipd = "A",
  trt_agd = "B",
  trt_common = NULL,
  endpoint_name = "Overall Survival",
  time_scale = "week",
  zph_transform = "log",
  zph_log_hazard = TRUE
)
```

## Anchored case

Similar tests can be done for the anchored case.

```{r}
data(weighted_twt)
data(adtte_twt)
data(pseudo_ipd_twt)

ph_diagplot(
  weights_object = weighted_twt,
  tte_ipd = adtte_twt,
  tte_pseudo_ipd = pseudo_ipd_twt,
  trt_var_ipd = "ARM",
  trt_var_agd = "ARM",
  trt_ipd = "A",
  trt_agd = "B",
  trt_common = "C",
  endpoint_name = "Overall Survival",
  time_scale = "week",
  zph_transform = "log",
  zph_log_hazard = TRUE
)
```


```{r, echo = FALSE, eval = FALSE}
# heuristic check
library(dplyr)
library(survival)
library(survminer)

ipd_matched <- weighted_sat$data
combined_data_tte <- ipd_matched %>%
  left_join(adtte_sat, by = c("USUBJID", "ARM"))
pseudo_ipd <- pseudo_ipd_sat
pseudo_ipd$weights <- 1

combined_data_tte <- rbind(
  combined_data_tte[, colnames(pseudo_ipd)],
  pseudo_ipd
)

# Change the reference treatment to B
combined_data_tte$ARM <- stats::relevel(as.factor(combined_data_tte$ARM), ref = "B")

combined_data_tte$TIME <- combined_data_tte$TIME / 30.4375 # change time to months

fit_surv <- survfit(Surv(TIME, EVENT == 1) ~ ARM, weights = weights, data = combined_data_tte)
fit_cox <- coxph(Surv(TIME, EVENT == 1) ~ ARM, weights = weights, data = combined_data_tte)

ggsurvplot(fit_surv,
  data = combined_data_tte,
  xlab = "Months",
  ylab = "Log-log plot OS",
  main = "Log-log plot OS",
  fun = "cloglog"
)

test.ph <- cox.zph(fit_cox, transform = "log")
print(ggcoxzph(test.ph))
```

# References
