---
title: "Unanchored Binary Analysis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: references.bib
csl: biomedicine.csl
vignette: >
  %\VignetteIndexEntry{Unanchored Binary Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Loading R packages

```{r}
# install.packages("maicplus")
library(maicplus)
```

Additional R packages for this vignette:

```{r}
library(dplyr)
```

# Illustration using example data

This example reads in `centered_ipd_sat` data that was created in `calculating_weights` vignette and uses `adrs_sat` dataset to run binary outcome analysis using the `maic_unanchored` function by specifying `endpoint_type = "binary"`.

Note that parameters `ipd` and `pseudo_ipd` in the `maic_unanchored` function for binary data analysis needs to have the following columns: USUBJID, ARM, RESPONSE. USUBJID in `ipd` needs to match USUBJID in `weights_object`. `pseudo_ipd` for binary data can be conveniently generated using `get_pseudo_ipd_binary` function.

Robust standard error for the adjusted result are calculated by sandwich variance estimator in `sandwich` package with the function `vcovHC`. Default type of variance estimator (specified by parameter `binary_robust_cov_type`) is `HC3`, but other types can be specified. For more information on different types, see `vcovHC`.

```{r}
data(centered_ipd_sat)
data(adrs_sat)

centered_colnames <- c("AGE", "AGE_SQUARED", "SEX_MALE", "ECOG0", "SMOKE", "N_PR_THER_MEDIAN")
centered_colnames <- paste0(centered_colnames, "_CENTERED")

weighted_data <- estimate_weights(
  data = centered_ipd_sat,
  centered_colnames = centered_colnames
)

# get dummy binary pseudo IPD
pseudo_adrs <- get_pseudo_ipd_binary(
  binary_agd = data.frame(
    ARM = "B",
    RESPONSE = c("YES", "NO"),
    COUNT = c(280, 120)
  ),
  format = "stacked"
)

result <- maic_unanchored(
  weights_object = weighted_data,
  ipd = adrs_sat,
  pseudo_ipd = pseudo_adrs,
  trt_ipd = "A",
  trt_agd = "B",
  normalize_weight = FALSE,
  endpoint_type = "binary",
  endpoint_name = "Binary Endpoint",
  eff_measure = "OR",
  # binary specific args
  binary_robust_cov_type = "HC3"
)
```

There are two summaries available in the result: descriptive and inferential. In the descriptive section, we have summaries of events.

```{r}
result$descriptive
```

In the inferential section, we have the fitted models stored (i.e. logistic regression) and the results from the `glm` models (i.e. odds ratios and CI). If other effect measures are needed besides odds ratios, we have an option to fit risk ratios or risk differences via `eff_measure` parameter. Here is the overall summary.

```{r}
result$inferential$summary
```

Here are model and results before adjustment.

```{r}
result$inferential$fit$model_before
result$inferential$fit$res_AB_unadj
```

Here are model and results after adjustment.

```{r}
result$inferential$fit$model_after
result$inferential$fit$res_AB
```

```{r, eval = FALSE, echo = FALSE}
# heuristic check
# merge in adrs with ipd_matched
ipd_matched <- weighted_data$data
combined_data_binary <- ipd_matched %>%
  left_join(adrs_sat, by = c("USUBJID", "ARM"))
pseudo_adrs$weights <- 1

combined_data_binary <- rbind(
  combined_data_binary[, colnames(pseudo_adrs)],
  pseudo_adrs
)

# Change the reference treatment to B
combined_data_binary$ARM <- stats::relevel(as.factor(combined_data_binary$ARM), ref = "B")

binobj_dat <- glm(RESPONSE ~ ARM, combined_data_binary, family = binomial(link = "logit"))
binobj_dat_adj <- glm(RESPONSE ~ ARM, combined_data_binary,
  weights = weights,
  family = binomial(link = "logit")
) |> suppressWarnings()

bin_robust_cov <- sandwich::vcovHC(binobj_dat_adj, type = "HC3")
bin_robust_coef <- lmtest::coeftest(binobj_dat_adj, vcov. = bin_robust_cov)
bin_robust_ci <- lmtest::coefci(binobj_dat_adj, vcov. = bin_robust_cov)

bin_robust_ci
exp(bin_robust_ci)

exp(summary(binobj_dat)$coef[2, "Estimate"])
exp(summary(binobj_dat_adj)$coef[2, "Estimate"])
```

# Using bootstrap to calculate standard errors

If bootstrap standard errors are preferred, we need to specify the number of bootstrap iteration (`n_boot_iteration`) in `estimate_weights` function and proceed fitting `maic_unanchored` function. Then, the outputs include bootstrapped CI. Different types of bootstrap CI can be found by using parameter `boot_ci_type`. See `boot.ci` in `boot` package for more details.

```{r}
weighted_data2 <- estimate_weights(
  data = centered_ipd_sat,
  centered_colnames = centered_colnames,
  n_boot_iteration = 100,
  set_seed_boot = 1234
)

result_boot <- maic_unanchored(
  weights_object = weighted_data2,
  ipd = adrs_sat,
  pseudo_ipd = pseudo_adrs,
  trt_ipd = "A",
  trt_agd = "B",
  normalize_weight = FALSE,
  endpoint_type = "binary",
  endpoint_name = "Binary Endpoint",
  eff_measure = "OR",
  boot_ci_type = "perc",
  # binary specific args
  binary_robust_cov_type = "HC3"
)

result_boot$inferential$fit$boot_res_AB
```
