---
title: "Bootstrap design"
output:
  html_document:
    code_folding: hide
    theme: spacelab 
    toc: true
    number_sections: true
---


<style type="text/css">

body{ /* Normal  */
      font-size: 14px;
  }
td {  /* Table  */
    font-size: 14px;
}
h1.title {
  font-size: 28px;
  color: DarkRed;
}
h1 { /* Header 1 */
    font-size: 22px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 18px;
  color: DarkBlue;
}
h3 { /* Header 3 */
    font-size: 18px;
  color: DarkBlue;
}

h4 { /* Header 4 */
    font-size: 18px;
  color: DarkBlue;
}

code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>



# Method 1 

## Steps
- Create bootstrap_HR() function to get the bootstrap weighted hazard ratios.
- The bootstrap weighted hazard ratios will then be used as an input of the standard boot() function in order to obtain the
bootstrap confidence interval of the weighted HR. 
- The boot() function will be included in the maic_tte_unanchor() function of the maic_tte_unanchored.R file so that we can get 
the results of all the different approaches (i.e., sandwich estimator, non-parametric bootstrap, posson approximation) for the unanchored cases
of time to event endpoints. 
- Add additional input parameter in the maic_tte_unanchor() function to meet the requirements for the boot() function
- Report bootstrap results 



## Bootstrapping for MAIC weighted hazard ratios 
  
The bootstrap_HR() function will be created to get the bootstrap weighted hazard ratios.
This function will include the following input parameters:
  
- *intervention_data:* A data frame containing individual patient data from the internal IPD study including
columns (treatment, time, status, centered baseline characteristics)
- *centered_colnames:* A character vector giving the names of the covariates to use in matching. 
These names must match the column names in intervention_data.
- *i:* Index used to select a sample within the boot() function.
- *model:* A model formula in the form 'Surv(time, status == 1) ~ treatment '. 
Variable names need to match the corresponding columns in intervention_data.
- *comparator_data:* A data frame containing pseudo individual patient data
from the comparator study needed to derive the relative treatment effect
including columns (time, event, weight=1, scaled_weight=1).
The outcome variables names must match intervention_data.
- *min_weight:* A numeric value that defines the minimum weight allowed.
This value (default 0.0001) will replace weights estimated at 0 in a sample.
- *trt_ext:* A character giving the name of the comparator treatment to be used as reference


Note: The bootstrap_HR() is intended to be used in conjunction with the boot() function to return the statistic 
to be bootstrapped. In this case by performing MAIC weighting using estimate_weights() function and returning a 
weighted hazard ratio (HR) from a Cox proportional hazards model. This is used as the 'statistic' argument in
the boot function.

From the bootstrap_HR() function we will export the HR as a numeric value.  


### Example code 

```{r}
bootstrap_HR <- function(intervention_data, centered_colnames, i, model, comparator_data, min_weight = 0.0001, trt_ext) {
  # create a visible binding for R CMD check
  wt <- NULL
  
  # Samples the data
  bootstrap_data <- intervention_data[i, ]
  
  # Estimates weights
  perform_wt <- estimate_weights(data = bootstrap_data, centered_colnames = centered_colnames)
  
  
  # Give comparator data weights of 1
  comparator_data_wts <- comparator_data %>% dplyr::mutate(weights = 1, scaled_weights = 1)
  
  # Add the comparator data
  combined_data <- dplyr::bind_rows(perform_wt$data, comparator_data_wts)
  combined_data$treatment <- stats::relevel(as.factor(combined_data$treatment), ref = trt_ext)
  
  # set weights that are below min_weight to min_weight to avoid issues with 0 values
  combined_data$wt <- ifelse(combined_data$weights < min_weight, min_weight, combined_data$weights)
  
  # survival data stat
  cox_model <- survival::coxph(model, data = combined_data, weights = wt)
  HR <- exp(cox_model$coefficients)
} 
  
```

## Bootstrap the confidence interval of the weighted HR 
  
The boot() function will include the following in the same format as defined in the bootstrap_HR() function: 

- Intervention data
- The statistic to be bootsrapped; i.e. bootstap_HR
- Number of bootstrap sampes
- Comparator pseudo data
- Matching variables
- Model to fit 


  
### Example code

```{r}
HR_bootstraps <- boot(
    data = dat_centered, # intervention data
    statistic = bootstrap_HR, # bootstrap the HR 
    R = 1000, # number of bootstrap samples
    comparator_data = dat_ext, # comparator pseudo data
    centered_colnames = grep("_CENTERED$", names(dat_centered)), # matching variables
    model = Surv(time, status == 1) ~ treatment # model to fit
  )

```


## Additional input parameter in the maic_tte_unanchor() function

In the maic_tte_unanchor() function of the maic_tte_unanchored.R file we need to add an additional input parameter as this function will
use the boot() function. Specifically, the dat_centered which is the intervention data as defined in the bootstrap_HR() function will be added. 


## Reporting bootstrap results 

Report the following: 

- Summarize bootstrap estimates in a histogram
- Median of the bootstrap samples
- Percentile CIs (This method takes the 2.5th and 97.5th percentiles)
- Bias-corrected and accelerated (BCa) CI (This method attempts to correct for any bias and skewness in the distribution of bootstrap estimates)

Add section 'Report 4: Bootstrapping' in the maic_tte_unanchor() function of the maic_tte_unanchored.R file


### Example code

```{r}
  
  # ==> Report 4: Bootstrapping
  
  ## Bootstrapping diagnostics
  # Summarize bootstrap estimates in a histogram
  # Vertical lines indicate the median and upper and lower CIs
  hist(HR_bootstraps$t, main = "", xlab = "Boostrapped HR")
  abline(v = quantile(HR_bootstraps$t, probs = c(0.025, 0.5, 0.975)), lty = 2)

  # Median of the bootstrap samples
  HR_median <- median(HR_bootstraps$t)

  # Bootstrap CI - Percentile CI
  boot_ci_HR <- boot.ci(boot.out = HR_bootstraps, index = 1, type = "perc")

  # Bootstrap CI - BCa CI
  boot_ci_HR_BCA <- boot.ci(boot.out = HR_bootstraps, index = 1, type = "bca")

  res[["boot_hist"]] <- grDevices::recordPlot()
  res[["HR_median"]] <- HR_median
  res[["boot_ci_HR"]] <- boot_ci_HR
  res[["boot_ci_HR_BCA"]] <- boot_ci_HR_BCA`

```


# Another way

```{r}
#' Bootstrapping weighted hazard ratios to obtain confidence interval
#'
#' This function uses [\code{\link{boot}}] to resample internal IPD data
#' to obtain bootstrapped hazard ratio values which is used to form confidence
#' interval. Note that due to lack of full IPD of covariates in the 
#' aggregate data, pseudo IPD will not be bootstrapped.
#' 
#' @param ipd_centered A data frame containing individual patient data
#'   from the internal IPD study. This data frame should have already been
#'   centered using [center_ipd] function
#' @param i Index used to select a sample within \code{\link{boot}}.
#' @param centered_colnames A character vector giving the names of the covariates to use in matching. These names must match the column names in internal dataset.
#' @param psuedo_ipd A data frame containing pseudo individual patient data
#'   from the comparator study needed to derive the relative treatment effect.
#'   Pseudo IPD is only used for unanchored case. pseudo_ipd should have time, event,
#'   and ARM variable with a same name and order as that of ipd_centered
#' @param model Survival model to run to calculate HR with correct time, event, and #'   ARM names
#' @param ref_treat Reference treatment for the survival model
#' @param min_weight A numeric value that defines the minimum weight allowed. 
#'   This value (default 0.0001) will replace weights estimated at 0 in a sample.
#'
#' @return The HR bootstraps
#' @export

bootstrap_HR <- function(ipd_centered, i, centered_colnames, 
                         pseudo_ipd = NULL, model,
                         min_weight = 0.0001, ref_treat = NULL){
  
  # Resamples the centered internal IPD data
  bootstrap_data <- ipd_centered[i,]
  
  # Estimates weights
  weighted_data <- estimate_weights(data = bootstrap_data, 
                                centered_colnames = centered_colnames)
  ipd_matched <- weighted_data$data
  
  if(length(unique(ipd_matched$ARM)) == 2){ #this is the anchored case
    combined_data <- ipd_matched
  } else if(length(unique(ipd_matched$ARM)) == 1){ #this is the unanchored case
    
    if(is.null(pseudo_ipd)) stop("For unanchored case pseudo_ipd needs to be specified")
    if(is.null(pseudo_ipd$ARM)) stop("external data should have an ARM column")
    
    # Assign weight of 1 for pseudo_ipd if it not defined already
    pseudo_ipd$weights <- 1

    combined_data <- rbind(ipd_matched[,colnames(pseudo_ipd)],
                          pseudo_ipd)
  } else {
    stop("ipd_matched ARM should have 1 or 2 treatments")
  }
    
  # set the base treatment
  if(is.null(ref_treat)) stop("Need to specify a value for ref_treat")
  combined_data$ARM <- stats::relevel(as.factor(combined_data$ARM), ref = ref_treat)
  
  # set weights that are below min_weight to min_weight to avoid issues with 0 values
  combined_data$weights <- ifelse(combined_data$weights < min_weight, min_weight, combined_data$weights)

  cox_model <- survival::coxph(model, data = combined_data, weights = weights, robust = TRUE)
  HR <- exp(cox_model$coefficients)

  return(HR)
}
```


```{r}
library(dplyr)
library(boot)

#set directory
setwd("~/GitHub/maicplus")
devtools::load_all()

adsl <- read.csv(system.file("extdata", "adsl.csv", package = "maicplus",
                             mustWork = TRUE))
adrs <- read.csv(system.file("extdata", "adrs.csv", package = "maicplus",
                             mustWork = TRUE))
adtte <- read.csv(system.file("extdata", "adtte.csv", package = "maicplus", mustWork = TRUE))

# Data containing the matching variables
adsl <- adsl %>%
  mutate(SEX_MALE = ifelse(SEX=="Male", 1, 0)) %>%
  mutate(AGE_SQUARED = AGE^2)

# Could use built-in function for dummizing variables 
# adsl <- dummize_ipd(adsl, dummize_cols=c("SEX"), dummize_ref_level=c("Female"))

# Response data  
adrs <- adrs %>%
  transmute(USUBJID, ARM, RESPONSE=AVAL)

# Time to event data (overall survival)
adtte <- adtte %>% 
  mutate(EVENT = 1-CNSR) %>% 
  transmute(USUBJID, ARM, TIME = AVAL, EVENT)

# Combine all ipd data
ipd <- adsl %>%
  full_join(adrs, by = c("USUBJID", "ARM")) %>%
  full_join(adtte, by = c("USUBJID", "ARM"))
head(ipd)

agd <- data.frame(
  STUDY = "Lung study",
  ARM = "Total",
  N = 300,
  AGE_MEAN = 51,
  AGE_MEDIAN = 49,
  AGE_SD = 3.25,
  SEX_MALE_PROP = 147/300,
  ECOG0_PROP = 0.40,
  SMOKE_PROP = 58/(300-5),
  N_PR_THER_MEDIAN = 2
)

#### prepare data
ipd_centered <- center_ipd(ipd = ipd, agd = agd)
head(ipd_centered)

centered_colnames <- c("AGE", "AGE_SQUARED", "SEX_MALE", "ECOG0", "SMOKE", "N_PR_THER_MEDIAN")
centered_colnames <- paste0(centered_colnames, "_CENTERED")
weighted_data <- estimate_weights(data = ipd_centered, 
                              centered_colnames = centered_colnames)

pseudo_ipd <- read.csv(system.file("extdata", "psuedo_IPD.csv", package = "maicplus", mustWork = TRUE))
ipd_matched <- weighted_data$data
pseudo_ipd$ARM <- "B"

# Need to ensure pseudo_ipd and ipd_matched have same names
colnames(pseudo_ipd) <- c("TIME", "EVENT", "ARM")

HR_bootstraps <- boot(data = ipd_matched,
  statistic = bootstrap_HR, 
  centered_colnames = centered_colnames, 
  pseudo_ipd = pseudo_ipd,
  model = Surv(TIME, EVENT==1) ~ ARM,
  ref_treat = "B",
  R = 1000)

# Median of the bootstrap samples
HR_median <- median(HR_bootstraps$t)

# Bootstrap CI - Percentile CI
boot_ci_HR <- boot.ci(boot.out = HR_bootstraps, index=1, type="perc")

# Bootstrap CI - BCa CI
boot_ci_HR_BCA <- boot.ci(boot.out = HR_bootstraps, index=1, type="bca")

HR_median
boot_ci_HR
boot_ci_HR_BCA
```

