---
title: "Forest Plot design to present unadjusted and adjusted results"
output:
  html_document:
    code_folding: hide
    theme: spacelab 
    toc: true
    number_sections: true
---

  
<style type="text/css">
  
body{ /* Normal  */
      font-size: 14px;
  }
td {  /* Table  */
    font-size: 14px;
}
h1.title {
  font-size: 28px;
  color: DarkRed;
}
h1 { /* Header 1 */
    font-size: 22px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 18px;
  color: DarkBlue;
}
h3 { /* Header 3 */
    font-size: 18px;
  color: DarkBlue;
}

h4 { /* Header 4 */
    font-size: 18px;
  color: DarkBlue;
}

code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>
  
  
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 6
)
``` 


# Load libraries
```{r}
library(gridExtra)
library(ggplot2)
```

# Background 

Design forest plot to visualize the direction and magnitude of the unadjusted and adjusted results for each category of endpoints.
We would have a separate forest plot including all the time to event endpoints, binary endpoints etc.

In this example I have used the overall survival and more endpoints could be added (e.g. PFS etc.) in the same plot.
An error bar with a diamond will be presented to indicate the point estimate with the corresponding 95% confidence interval for the OS unadjusted and adjusted results. Next to the forest plot and in the same line with the error bars we will present two columns that will show the coresponding HR [95% CI] and corresponding p-value. 


# Steps

- Load data 
- Modify the report_table function in the matching_unanchored_tte.R file to get the results with a separate column for the HR and the 95% CI
- In the maic_tte_unanchor function of the matching_unanchored_tte.R file, call the modified report_table so that the modified results are printed
- Run the modified maic_tte_unanchor function to get the unadjusted and adjusted results 
- Use the results to create a forest plot 


## Load Data
```{r}
devtools::load_all()
#### load data ----------------------------------------------------------

### IPD
# Read in relevant ADaM data and rename variables of interest
adsl <- read.csv(system.file("extdata", "adsl.csv",
  package = "maicplus",
  mustWork = TRUE
))
adrs <- read.csv(system.file("extdata", "adrs.csv",
  package = "maicplus",
  mustWork = TRUE
))
adtte <- read.csv(system.file("extdata", "adtte.csv",
  package = "maicplus",
  mustWork = TRUE
))

### AgD
# Baseline aggregate data for the comparator population
target_pop <- read.csv(system.file("extdata", "aggregate_data_example_1.csv",
  package = "maicplus", mustWork = TRUE
))


# for time-to-event endpoints, pseudo IPD from digitalized KM
pseudo_ipd <- read.csv(system.file("extdata", "psuedo_IPD.csv",
  package = "maicplus",
  mustWork = TRUE
))

#### prepare data ----------------------------------------------------------
target_pop <- process_agd(target_pop)
adsl <- dummize_ipd(adsl, dummize_cols = c("SEX"), dummize_ref_level = c("Female"))
use_adsl <- center_ipd(ipd = adsl, agd = target_pop)

match_res <- estimate_weights(
  data = use_adsl,
  centered_colnames = grep("_CENTERED$", names(use_adsl)),
  start_val = 0,
  method = "BFGS"
)
```



## Add report_table_ff in the matching_unanchored_tte.R file to separate HR from CI 
```{r}
##### ADDED
report_table_ff <- function(coxobj, medSurvobj, tag = NULL) {
  hr_res <- format(round(summary(coxobj)$conf.int[-2], 2), nsmall = 2)
  # hr_res <- paste0(hr_res[1], "[", hr_res[2], ";", hr_res[3], "]")      #### DELETED
  hr_pval <- format(round(summary(coxobj)$waldtest[3], 3), nsmall = 3)
  if (hr_pval == "0.000") hr_pval <- "<0.001"

  meds_report <- format(round(medSurvobj[, c("median", "0.95LCL", "0.95UCL")], 1), nsmall = 1)
  meds_report <- apply(meds_report, 1, function(xx) paste0(xx[1], "[", xx[2], ";", xx[3], "]"))

  desc_res <- cbind(medSurvobj[, "treatment", drop = F],
    data.frame(N = round(medSurvobj$n.max, 1)),
    "n.events(%)" = paste0(
      round(medSurvobj$events, 1), "(",
      format(round(medSurvobj$events * 100 / medSurvobj$n.max, 1), nsmall = 1), ")"
    ),
    "median[95% CI]" = meds_report
  )

  desc_res <- cbind(desc_res[c(2, 1), ], "HR" = hr_res[1], "lowerCI" = hr_res[2], "upperCI" = hr_res[3], "WaldTest" = c(hr_pval, "")) #### CHANGED

  if (!is.null(tag)) desc_res <- cbind(data.frame(Matching = rep(tag, nrow(desc_res))), desc_res)
  desc_res
}




maic_tte_unanchor <- function(useWt, dat, dat_ext, trt, trt_ext,
                              time_scale = "month", endpoint_name = "OS",
                              transform = "log") {
  timeUnit <- list("year" = 365.24, "month" = 30.4367, "week" = 7, "day" = 1)

  if (length(useWt) != nrow(dat)) stop("length of useWt should be the same as nrow(dat)")
  if (!time_scale %in% names(timeUnit)) stop("time_scale has to be 'year', 'month', 'week' or 'day'")

  res <- list()

  # set up IPD 'dat' with maic weights
  dat$weight <- useWt
  dat <- dat[, c("treatment", "time", "status", "weight")]

  # set up pseudo IPD 'dat_ext' with universal weight of 1
  dat_ext <- dat_ext[dat_ext$treatment == trt_ext, ]
  dat_ext$treatment <- trt_ext
  dat_ext$weight <- 1
  dat_ext <- dat_ext[, names(dat)]

  # merge pseudo IPD and real ipd
  dat <- rbind(dat, dat_ext)
  dat$treatment <- factor(dat$treatment, levels = c(trt_ext, trt))

  # ==> Report 1: KM plot

  # derive km w and w/o weights
  kmobj <- survfit(Surv(time, status) ~ treatment, dat, conf.type = "log-log")
  kmobj_adj <- survfit(Surv(time, status) ~ treatment, dat, weights = dat$weight, conf.type = "log-log")

  par(cex.main = 0.85)
  km_plot(kmobj, kmobj_adj,
    time_scale = time_scale,
    trt = trt, trt_ext = trt_ext,
    endpoint_name = endpoint_name
  )
  res[["plot_km"]] <- grDevices::recordPlot()

  res[["fit_km_data_before"]] <- survfit_makeup(kmobj)
  res[["fit_km_data_after"]] <- survfit_makeup(kmobj_adj)

  # ==> Report 2: Analysis table (Cox model) before and after matching, incl Median Survival Time

  # derive median survival time
  medSurv <- medSurv_makeup(kmobj, legend = "before matching", time_scale = time_scale)
  medSurv_adj <- medSurv_makeup(kmobj_adj, legend = "after matching", time_scale = time_scale)
  medSurv_out <- rbind(medSurv, medSurv_adj)

  res[["report_median_surv"]] <- medSurv_out

  # fit PH Cox regression model
  coxobj <- coxph(Surv(time, status) ~ treatment, dat, robust = T)
  coxobj_adj <- coxph(Surv(time, status) ~ treatment, dat, weights = dat$weight, robust = T)

  res[["fit_cox_model_before"]] <- coxobj
  res[["fit_cox_model_after"]] <- coxobj_adj

  res[["report_overall"]] <- rbind(
    report_table(coxobj, medSurv, tag = paste0("Before/", endpoint_name)),
    report_table(coxobj_adj, medSurv_adj, tag = paste0("After/", endpoint_name))
  )

  ########### ADDED
  res[["report_overall_ff"]] <- rbind(
    report_table_ff(coxobj, medSurv, tag = paste0("Before/", endpoint_name)),
    report_table_ff(coxobj_adj, medSurv_adj, tag = paste0("After/", endpoint_name))
  )


  # output
  res
}
```



## Run maic_tte_unanchor example and get the data frame with the results
```{r, fig.show='hide' }
useWt <- as.vector(match_res$data[, "weights"])
adtte$weight <- useWt
dat <- adtte[, c("ARM", "AVAL", "EVNT")]
dat_centered <- cbind(dat, match_res$data[, match_res$centered_colnames])
colnames(dat) <- c("treatment", "time", "status")
dat_ext <- pseudo_ipd
dat_ext$ARM <- "B"
dat_ext <- dat_ext[, c("ARM", "Time", "Event")]
colnames(dat_ext) <- c("treatment", "time", "status")
trt <- "A"
trt_ext <- "B"


tte_unanchor <- maic_tte_unanchor(useWt, dat, dat_ext, trt, trt_ext, time_scale = "month", endpoint_name = "OS", transform = "log")


d <- tte_unanchor$report_overall_ff[c(1, 3), c(1, 6, 7, 8, 9)]
```



## Create forest plot

```{r}
# Create forest plot

d$Index <- 1:dim(d)[1]

# Transform HR and CIs to number values
d$HR <- as.numeric(d$HR)
d$lowerCI <- as.numeric(d$lowerCI)
d$upperCI <- as.numeric(d$upperCI)

plot1 <- ggplot(d, aes(y = Index, x = HR)) +
  geom_point(shape = 18, size = 4) +
  geom_errorbarh(aes(xmin = lowerCI, xmax = upperCI), height = 0.1) +
  geom_vline(xintercept = 1, color = "red", linetype = "dashed", cex = 1, alpha = 0.5) +
  coord_cartesian(xlim = c(0, 1.5), ylim = c(0, 5)) +
  scale_y_continuous(name = "", breaks = 1:2, labels = d$Matching, trans = "reverse") +
  xlab("Hazard Ratio (95% CI)") +
  ylab(" ") +
  theme_bw() +
  theme(
    panel.border = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    axis.text.y = element_text(size = 12, colour = "black"),
    axis.text.x.bottom = element_text(size = 12, colour = "black"),
    axis.title.x = element_text(size = 12, colour = "black")
  )

# Create the table-base pallete
table_base <- ggplot(d, aes(y = Matching)) +
  coord_cartesian(xlim = c(-10, 10), ylim = c(0, 5)) +
  ylab(NULL) +
  xlab("  ") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 12),
    axis.text.x = element_text(color = "white", hjust = -3, size = 25), ## This is used to help with alignment
    axis.line = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none",
    panel.background = element_blank(),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank()
  )




## HR 95% CI table
tab1 <- table_base +
  geom_text(aes(y = rev(Index), x = 1, label = paste0(d[, 2], "[", d[, 3], ";", d[, 4], "]")), size = 4) +
  ggtitle("HR[95% CI]")

## pvalue table
tab2 <- table_base +
  geom_text(aes(y = rev(Index), x = 1, label = WaldTest), size = 4) +
  ggtitle("pvalue")

## Merge tables with plot
lay <- matrix(c(1, 1, 1, 2, 3), nrow = 1)
grid.arrange(plot1, tab1, tab2, layout_matrix = lay)
```



## Alternative Forest Plot 

Another example: Figure 4 of [this publication](https://pubmed.ncbi.nlm.nih.gov/35764490/)



